---
title: "ardor"
author: "Anis Mansouri"
date: "11 January 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Packages:
library(minet)
library(viper)
library(vsn)
library(infotheo)
library(igraph)
#library(org.Hs.eg.db)
#library(EnsDb.Hsapiens.v86)
#library(biomaRt)
#library(Rgraphviz)
```

# Loading the data

```{r}
# Raw data:
raw_data= read.table("GSE119931_PANC1.FOXA2KO.genes.counts.txt", header = TRUE)

# Extract the counts data and renames its rows:
raw_counts=raw_data[,c(7:12)]
rownames(raw_counts)= raw_data$Geneid
```

# Data cleaning

```{r}
# Remove genes with 0 expression level:
raw_counts= raw_counts[rowSums(raw_counts) > 0,]
```

# Data normalization

```{r}
vsn_parameters=vsnMatrix(as.matrix(raw_counts))
normalized_counts = as.data.frame(vsn::predict(vsn_parameters, as.matrix(raw_counts)))
```

# Creation of a transposed counts data set

```{r}
# build.mim() resuires a transposed data set (samples in rows and genes in columns):
normalized_counts_transposed= as.data.frame(t(as.matrix(normalized_counts)))
```

# Data discretization (required when a discrete entropy estimator is used)

```{r}
discretized_counts=discretize(normalized_counts_transposed, nbins = 10)
rownames(discretized_counts)=rownames(normalized_counts_transposed)
```

# Building the mutual information matrices (MIM)

```{r}
# Mutual Information Matrix for continuous counts:
mim_continuous=build.mim(normalized_counts_transposed)

# Mutual Information Matrix for discrete counts:
mim_doscrete=build.mim(discretized_counts, estimator = "mi.empirical")  
```

# Network inference by ARACNE

```{r}
# Infer the network using the continuous counts:
network_1= aracne(mim_continuous)
network_1[is.na(network_1)]=0

# Infer the network using the discrete counts:
network_2= aracne(mim_doscrete)
network_2[is.na(network_2)]=0
```

# Creation of the adjacency file (3 columns: tf, target, weight)

```{r}
# Get the significant gene-gene interactions:
pairwise_interactions_1=graph.adjacency(network_1, mode = 'upper', weighted = TRUE)
pairwise_interactions_2=graph.adjacency(network_2, mode = 'upper', weighted = TRUE)

# Write the adjacency files:
write.table(get.data.frame(pairwise_interactions_1), "adjacency_1.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(get.data.frame(pairwise_interactions_2), "adjacency_2.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

# Getting the regulons from the inferred networks

```{r}
# Regulons from networks built using continuous counts:
regulons_1 = aracne2regulon(afile = "adjacency_1.txt", eset = as.matrix(normalized_counts), format = "3col", verbose = FALSE)

# Regulons from networks built using discrete counts:
regulons_2 = aracne2regulon(afile = "adjacency_2.txt", eset = as.matrix(discretized_counts), format = "3col", verbose = FALSE)
```

# Estimation of the differential activity of regulatory proteins between PANC1.WT and PANC1.FOXA2.KO (in progress)

```{r}
# Generating the regulatory protein activity matrix:
normalized_counts$gene=rownames(normalized_counts)
normalized_counts=normalized_counts[, c(7, 1:6)]
rownames(normalized_counts)= str_replace(rownames(normalized_counts),
                        pattern = ".[0-9]+$",
                        replacement = "")

protein_activity=viper(eset=normalized_counts, regulon = regulons_1, minsize = 5, eset.filter = FALSE,cores = 1, 
                                       verbose = FALSE, method = c("scale"))
```

# Gene mapping (ignore this chunk)

```{r}
library(stringr)
gene_ids <- str_replace(raw_data$Geneid, çpattern = ".[0-9]+$",
                        replacement = "")
a=AnnotationDbi::select(org.Hs.eg.db,gene_ids, "SYMBOL", "ENSEMBL")# 5755 NA
```


