# Libraries loading

```{r}
library(viper)
library(dorothea)
library(Homo.sapiens)
library(stringr)
library(binilib)
```

# Extraction of ABC-confidence interactions from Dorothea

```{r}
dorothea_abc_regulons=as.data.frame(dorothea_hs[dorothea_hs$confidence %in% c("A", "B", "C"), c(1,3)])
```
# Inference of the regulons from ARACNE adjacency matrix (aracne_adj.txt)

```{r}
# Load the processed counts dataset used to infer the interactome (norm_counts.csv: 10209 genes x 6 samples):
counts_set=read.csv("norm_counts.csv")
rownames(counts_set)=counts_set$X
counts_set$X=NULL

# Infer the regulons:
aracne_reg=aracne2regulon("aracne_adj.txt", eset = as.matrix(counts_set), format = "3col") # The Mode of Regulation (MoR) between each gene pair 
# of the adjacency file will be equal to the Spearman correlation coefficient between these genes. 
# The likelihood of the interaction will be equal to: MI(gene pair)/max(MI).

# Convert the regulon object into a dataframe:
regulons_df=as.data.frame(reg2tibble(aracne_reg))
```

# Conversion of the ENSEMBL IDs into SYMBOLS

```{r}
# Create the conversion function::
ensembl_into_symbol=function(x) {mapIds(Homo.sapiens,x, "SYMBOL", "ENSEMBL", multiVals = "first")}

# Remove the .version from the IDs:
regulons_df[,1:2]=str_replace(c(regulons_df$source, regulons_df$target), pattern = ".[0-9]+$", replacement = "")

# Convert into SYMBOLS IDs:
symbols_dataframe=as.data.frame(apply(regulons_df[c(1,2)], 2, ensembl_into_symbol))

# Remove genes whithout SYMBOLS IDs:
regulons_df_sym=na.omit(cbind(symbols_dataframe,regulons_df[-c(1:2)]))
rownames(regulons_df_sym)=NULL
```

# Finding the matching interactions between Dorothea and ARACNE regulons

## The function to check matching rows

```{r}
check.match= function(dataframe1, columns1, dataframe2, columns2) {
  
  df1_interactions = paste(dataframe1[[columns1[1]]], dataframe1[[columns1[2]]])
  df2_interactions = paste(dataframe2[[columns2[1]]], dataframe2[[columns2[2]]])
  df2_rev_interactions = paste(dataframe2[[columns2[2]]], dataframe2[[columns2[1]]])
  
  match_direction1=match(df1_interactions, df2_interactions,nomatch = 0)
  to_skip=which(match_direction1 !=0)
  match_direction2=match(df1_interactions[-c(to_skip)], df2_rev_interactions, nomatch = 0)

  match_direction1[which(match_direction1==0)]=match_direction2
  match_df1=which(match_direction1!=0)
  match_df2=match_direction1[which(match_direction1!=0)]
 
  cat(length(match_df1),"matching rows","\n","$df1_rows: matching rows from", 
  substitute(dataframe1),"\n", "$df2_rows: matching rows from", substitute(dataframe2),"\n")
  
  return(list("df1_rows"= match_df1, "df2_rows"=match_df2))
}
```
## Extraction of matching interactions between Dorothea and ARACNE regulons

```{r}
matching_rows=check.match(regulons_df_sym,c(1,2),dorothea_abc_regulons,c(1,2))
doro_int_aracne=dorothea_abc_regulons[matching_rows,]

# To match against the entire dorothea's regulons:
dorothea_regulons=as.data.frame(dorothea_hs[,c(1,3)])
matching_rows_all=check.match(regulons_df_sym,c(1,2),dorothea_regulons,c(1,2))
doro_all_int_aracne=dorothea_regulons[matching_rows_all,]

```
