---
path_to_wd <- '/Users/Aurelien/Dropbox/network-inference-by-aracne-in-R/'

knitr::opts_knit$set(root.dir = path_to_wd)
setwd(path_to_wd)
getwd()
```

```{r}
# Packages:
library(minet)
library(viper)
library(vsn)
library(infotheo)
library(igraph)

library(ggplot2)
library(reshape2)
```

# Loading the data

```{r}
# Raw data:
raw_data= read.table("GSE119931_PANC1.FOXA2KO.genes.counts.txt", header = TRUE)

# Extract the counts data and renames its rows:
raw_counts=raw_data[,c(7:12)]
rownames(raw_counts)= raw_data$Geneid
```

# Data cleaning

```{r}
# Remove genes with 0 expression level:
raw_counts= raw_counts[rowSums(raw_counts) > 0,]
```

```{r}
# Check data distribution and filter out low counts (noise)
raw_counts_long <- melt(log2(raw_counts))

ggplot(raw_counts_long, aes(x = variable, y = value, group = variable)) + geom_violin() 

raw_counts[log2(raw_counts) < 5] <- NA

raw_counts <- raw_counts[rowSums(is.na(raw_counts)) < dim(raw_counts)[2],]

raw_counts <- raw_counts[complete.cases(raw_counts),] #build.min cannot handle NA values
```


# Data normalization

```{r}
vsn_parameters=vsnMatrix(as.matrix(raw_counts))
normalized_counts = as.data.frame(vsn::predict(vsn_parameters, as.matrix(raw_counts)))
```

# Creation of a transposed counts data set

```{r}
# build.mim() resuires a transposed data set (samples in rows and genes in columns):
normalized_counts_transposed= as.data.frame(t(as.matrix(normalized_counts)))
```

# Data discretization (required when a discrete entropy estimator is used)

```{r}
discretized_counts=discretize(normalized_counts_transposed, nbins = 10)
rownames(discretized_counts)=rownames(normalized_counts_transposed)
```

# Building the mutual information matrices (MIM)

```{r}
# Mutual Information Matrix for continuous counts:
mim_continuous=build.mim(normalized_counts_transposed)

# Mutual Information Matrix for discrete counts:
mim_doscrete=build.mim(discretized_counts, estimator = "mi.empirical")  
```

# Network inference by ARACNE

```{r}
# Infer the network using the continuous counts:
network_1= aracne(mim_continuous)
network_1[is.na(network_1)]=0

plot(hist(network_1, breaks = 1000))

# Infer the network using the discrete counts:
network_2= aracne(mim_doscrete)
network_2[is.na(network_2)]=0

# save.image(file = "network_infered.RData")
```

# Creation of the adjacency file (3 columns: tf, target, weight)

```{r}
# Get the significant gene-gene interactions:
pairwise_interactions_1=graph.adjacency(network_1, mode = 'upper', weighted = TRUE)
pairwise_interactions_2=graph.adjacency(network_2, mode = 'upper', weighted = TRUE)

# Write the adjacency files:
write.table(get.data.frame(pairwise_interactions_1), "aracne_adj.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(get.data.frame(pairwise_interactions_2), "aracne_adj_disc.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

# Getting the regulons from the inferred networks

```{r}
# Regulons from networks built using continuous counts:
regulons_1 = aracne2regulon(afile = "aracne_adj.txt", eset = as.matrix(normalized_counts), format = "3col", verbose = FALSE)

# Regulons from networks built using discrete counts:
regulons_2 = aracne2regulon(afile = "aracne_adj_disc.txt", eset = as.matrix(discretized_counts), format = "3col", verbose = FALSE)
```

